---
title: "comparison"
author: "Roberto Sánchez"
date: "5/19/2022"
output: html_document
---

```{r}
library("tidyverse")
library("jsonlite")
library("RColorBrewer")
library("igraph")
library("ggraph")
library("cowplot")
library("readxl")
library("curl")
```

```{r}
source("notebook/parse_antismash.R")
dirs <- list.dirs(path = "F:/roberto_bgcs/bgcflow/processed", recursive = FALSE)
dirs <- dirs[str_detect(dirs, "et_al")]
dirs <- paste0(dirs, "/antismash/6.0.1/")
#data <- sapply(dirs, parse_study, simplify = FALSE)
data <- readRDS("data/comparison.rds")
```

```{r}
source("notebook/parse_bigscape.R")
bigscape_dir <- "F:/roberto_bgcs/bgcflow/interim/all_hqmags_study_bgcs_antismash_6.0.1"
bigscape <- import_bigscape(bigscape_dir)
```

```{r}
studies <- str_match(dirs, "\\/([\\w-]+?_et_al_\\d{4})")[,2]
data <- keep(data, map_lgl(data, ~is_empty(.x[["regions"]]) == FALSE))
names(data) <- studies

references <- c("Crits-Christoph_et_al_2018", 
             "Sharrar_et_al_2020", 
             "Chen_et_al_2020", 
             "Singleton_et_al_2021",
             "Liu_et_al_2021", 
             "Bickhart_et_al_2022")

environments <- c("Soil", 
                 "Soil", 
                 "Microbial Mats", 
                 "Activated Sludge", 
                 "Activated Sludge", 
                 "Sheep Feces")

long_reads <- c('short reads', 
                'short reads', 
                'short reads', 
                'long reads', 
                'long reads', 
                'long reads')

studies <- tibble(study_tag = LETTERS[1:length(studies)], 
       reference = references, 
       author = str_match(reference, "(^.+?)_et_al")[,2],
       environment = environments, 
       long_reads = long_reads) %>%
  mutate(author = factor(author, levels = author), 
         environment = str_replace(environment, " ", "\n"))

studies
```
```{r}
bgcs %>%
  mutate(product = map_chr(product, paste, collapse = ';')) %>%
  write_csv(file = "tables/comparison_bgcs.csv")
```


```{r}
traffic_light <- ggplot(studies, aes(x = author, label = environment)) +
  geom_point(y = 1, size = 5, aes(color = long_reads)) +
  geom_label(y = 0, size = 3) +
  scale_color_manual(values = c("steelblue","firebrick")) +
  scale_y_continuous(limits = c(-0.5,1.5), breaks = 0:1) +
  theme_minimal() +
  theme( 
    panel.grid.major.y = element_line(color = "black"), 
    panel.grid.minor = element_blank(), 
    legend.position = "top", 
    axis.text = element_text(color = "white" ), 
    legend.margin = margin(-1,-1,-1,-1)) +
  guides(color = guide_legend(title = NULL)) +
  ylab(" ") +
  xlab("")

traffic_light
```
Filtering Bickhart for HQ Bacterial MAGs

```{r}
Bickhart_supplementary <- 
  "https://static-content.springer.com/esm/art%3A10.1038%2Fs41587-021-01130-z/MediaObjects/41587_2021_1130_MOESM3_ESM.xlsx"

download.file(
  Bickhart_supplementary, 
  destfile = "data/bickhart_supplementary.xlsx", 
  quiet = TRUE,
  mode = "wb") 

bickhart_bacteria <- read_xlsx(
  "data/bickhart_supplementary.xlsx", 
  sheet = 'SupplementaryTable1') %>%
  setNames(c("mag", "tax")) %>%
  filter(str_detect(tax, "^d__Bacteria")) %>%
  pull(mag)

# Manual CheckM output

bickhart_hq <- read_tsv("data/bickhart_checkm.tsv", col_select = c(1,12,13)) %>%
  set_names("mag", "completeness", "contamination") %>%
  filter(completeness > 90) %>%
  filter(contamination < 5) %>%
  pull(mag) %>%
  str_remove("\\.contigs")
  
```

Filtering Liu

```{r}
Liu_supplementary <- 
  "https://static-content.springer.com/esm/art%3A10.1186%2Fs40168-021-01155-1/MediaObjects/40168_2021_1155_MOESM2_ESM.xlsx"

download.file(
  Liu_supplementary, 
  destfile = "data/liu_supplementary.xlsx", 
  quiet = TRUE,
  mode = "wb") 

Liu_df <- read_xlsx("data/liu_supplementary.xlsx", sheet = 3) %>%
  select(MAGs, `Completeness (%)`, `Contamination (%)`, `Taxa (GTDB)`) %>%
  set_names("mag", "completeness", "contamination", "tax") 

liu_bacteria <- Liu_df %>%
  filter(str_detect(tax, "^d__Bacteria")) %>%
  pull(mag)
  
liu_hq <- Liu_df %>%
  filter(completeness > 90) %>%
  filter(contamination < 5) %>%
  pull(mag)

```

```{r}
CritsChristoph_supplementary <-"https://static-content.springer.com/esm/art%3A10.1038%2Fs41586-018-0207-y/MediaObjects/41586_2018_207_MOESM3_ESM.xlsx"
  
download.file(
  CritsChristoph_supplementary, 
  destfile = "data/crits-christoph_supplementary.xlsx", 
  quiet = TRUE,
  mode = "wb") 

CritsChristoph_bioproject_accession <- 'PRJNA449266'
CritsChristoph_mapping <- paste0(
  "https://www.ncbi.nlm.nih.gov/bioproject/?term=", 
  CritsChristoph_bioproject_accession) %>%
  read_html() %>%
  html_node("#AssemblyDetails") %>%
  html_table() %>%
  select(Assembly, Isolate) %>%
  set_names(c("genome_id", "isolate")) %>%
  mutate(isolate = str_remove(isolate, '^.+? '))

critschristoph_hq <- read_xlsx("data/crits-christoph_supplementary.xlsx") %>%
  set_names(~tolower(str_replace_all(.x, "\\W", ''))) %>%
  mutate(genome = str_remove(genome, "^.+_")) %>%
  left_join(CritsChristoph_mapping, by = c('genome' = 'isolate')) %>%
  filter(checkmcompleteness > 90) %>%
  filter(checkmcontamination < 5) %>%
  pull(genome_id)
    
```

```{r}
Sharrar_supplementary <- 
  'https://journals.asm.org/doi/suppl/10.1128/mBio.00416-20/suppl_file/mbio.00416-20-st002.xlsx'

sharrar_genomes <- list.dirs(
  "../../bgcflow/processed/Sharrar_et_al_2020/antismash/6.0.1", 
  recursive = FALSE, 
  full.names = FALSE) 

download.file(Sharrar_supplementary,
              destfile = "data/sharrar_supplementary.xlsx",
              mode = "wb")

genome_mapping_pattern <- 
  paste0(sharrar_genomes, "$", collapse = "|")
  
sharrar_hq <- read_xlsx("data/sharrar_supplementary.xlsx") %>%
  set_names(~tolower(str_replace_all(.x, "\\W", ''))) %>%
  filter(checkmcompleteness > 90) %>%
  filter(checkmcontamination < 5) %>%
  mutate(corrected_genome_id = str_match(genomename, genome_mapping_pattern)[,1]) %>%
  drop_na(corrected_genome_id) %>%
  pull(corrected_genome_id) 
  
```


```{r}
# Manual checkm
chen_hq_bacteria <- read_tsv("data/chen_checkm_output.tsv") %>%
  set_names(~str_remove_all(.x, "\\W")) %>%
  filter(str_detect(Markerlineage, "rchaea") == FALSE) %>%
  filter(Completeness > 90) %>%
  filter(Contamination < 5) %>% 
  pull(BinId)
  

chen_api <- "https://api.mg-rast.org/project/mgp81948?verbosity=full"
chen_genome_ids <- fromJSON(chen_api)$metagenomes$metagenome_id
chen_urls <- paste0("https://api.mg-rast.org/download/", chen_genome_ids, "?file=299.1")
chen_destinations <- paste0('//wsl$/Ubuntu/home/roberto/chen/', chen_genome_ids, ".fna")
#map(1:length(chen_urls), ~download.file(chen_urls[.x], destfile =chen_destinations[.x]))
```

```{r}
bgcs <- map(data, "regions") %>%
  bind_rows(.id = "study") %>%
  left_join(studies[,c("reference", "author")], by = c("study" = "reference")) %>%
    # Some accession numbers in these two studies lost the ".1", 
    # so I add it again to be able to map it to BiG-SCAPE
    mutate(
      bgc_id = if_else(
        author %in% c("Singleton", "Crits-Christoph") &
          str_detect(bgc_id, "\\.1\\.region", negate = TRUE),
        str_replace(bgc_id, ".region", ".1.region"), bgc_id)) %>%
  left_join(bigscape, by = "bgc_id") %>%
  mutate(class = factor(translate_class(class), levels = unique(my_classes))) %>%
  filter(author != "Crits-Christoph" | genome_id %in% critschristoph_hq) %>%
  filter(author != "Sharrar" | genome_id %in% sharrar_hq) %>%
  filter(author != "Chen" | genome_id %in% chen_hq_bacteria) %>%
  filter(author != "Liu" | genome_id %in% liu_hq) %>%
  filter(author != "Bickhart" | genome_id %in% c(bickhart_hq, bickhart_bacteria))
  
```

# Contig edge comparison

```{r fig.width=9}
contig_edge_1 <- bgcs %>%
  arrange(author) %>%
  group_by(author) %>%
  mutate(perc = round(100 - 100*sum(contig_edge)/n(), 1), 
         label = paste0(author,"\n(", perc, "%)"), 
         label = factor(label, levels = unique(label))) %>%
  ungroup() %>%
  mutate(BGCs = if_else(contig_edge, "incomplete", "complete"), 
         BGCs = factor(BGCs, levels = c ("incomplete", "complete"))) %>%
  ggplot(aes(x = label, fill = BGCs)) +
  geom_bar(color = "black") +
  scale_fill_manual(values = c("grey95", "darkgrey")) +
  theme_minimal() +
  theme(axis.line = element_line())

contig_edge_2 <- contig_edge_1 +
  theme(axis.title = element_blank(),   
        legend.position = "top")

contig_edge_1

```

The short-read studies (first three) have BGCs mostly in the contig edge. 
The studies using long reads have mostly complete BGCs. 
Bickhart has the highest, but uses PacBio. 
Singleton and Liu use "hybrid assembly".

# Have different studies different classes of BGCs ?

```{r fig.width=9}
class_distribution_1 <- bgcs %>% 
  filter(is.na(class) == FALSE) %>%
  arrange(author) %>%
  group_by(author) %>%
  mutate(n_study = n(), 
         label = paste0(author, "\n(", n_study, " BGCs)"), 
         label = factor(label, levels = unique(label))) %>%
  ungroup() %>%
  group_by(label, class, n_study) %>%
  summarise(n = n(), perc = n / n_study * 100) %>%
  distinct() %>%
  ggplot(aes(x = class, y = perc, fill = class)) + 
  geom_bar(stat ="identity") +
  facet_grid(~label) +
  scale_fill_manual(values = my_colors) +
  #theme_minimal() +
  theme(axis.text.x = element_blank()) +
  ylab("% of BGCs") +
  xlab("") 

class_distribution_2 <- class_distribution_1 +
  theme(
    legend.margin = margin(-1,-1,-1,-1),
    #strip.text.x = element_blank(),
    legend.position = "bottom", 
    legend.direction = "horizontal", 
    axis.ticks = element_blank()) +
  guides(fill = guide_legend(nrow = 1, title = NULL))

class_distribution_1
```

Studies on soil have abundant NRPS (first two)
Bickhart (Sheep feces) has mostly RiPPs.

# What's the novelty of the BGCs in other studies?

```{r fig.width=9}
long_bigscape <- bigscape %>%
  select(-class) %>%
  left_join(bgcs[c("bgc_id", "author", "class", "product")], by = "bgc_id") %>%
  mutate(mibig = if_else(str_detect(bgc_id, "^BGC"), TRUE, FALSE)) %>%
  filter(is.na(author) == FALSE | mibig) %>%
  pivot_longer(starts_with('GCF'), names_to = "cutoff", values_to = "GCF") %>%
  mutate(cutoff = as.numeric(str_match(cutoff, "c(.+$)")[,2]))

# Possibly useful function "find_mibig_hits"

bgc_count <- table(bgcs$author)
bgc_count <- tibble(author = names(bgc_count), total_bgcs = bgc_count)

GCFs_with_mibig_hit <- long_bigscape %>%
  group_by(cutoff, GCF) %>%
  summarise(n_mibig = sum(mibig)) %>%
  filter(n_mibig > 0) %>%
  left_join(long_bigscape, by = c("cutoff", "GCF"))

novelty_1 <- GCFs_with_mibig_hit  %>%
  filter(mibig == FALSE) %>%
  left_join(bgc_count, by = "author") %>%
  group_by(author, cutoff, class) %>%
  summarise(hits = n(), 
            perc_hits = hits/total_bgcs * 100) %>%
  bind_rows(tibble(author = "Crits-Christoph")) %>%
  mutate(author = factor(author, levels = studies$author)) %>%
  ggplot(aes(x = cutoff, fill = class, y = perc_hits)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_grid(~author) +
  scale_fill_manual(values = my_colors, drop = FALSE) +
  ylab("% of BGCs clustered with a MIBiG BGC") 
  
novelty_1
  
novelty_2 <- novelty_1 + 
  theme(strip.text = element_text(angle = 90, hjust = 0.5),
        legend.position = "none")
```
Christoph has no hits at all, but the dataset is much smaller. 
Only Sharrar and Liu have hits at 0.3 

Our hits at c = 0.5 are:
arylpolyenes hits to flexirubin (BGC0000839)
ectoines hits to reference ectoines
NRPS-like hit to BGC0000871

# Do our BGCs cluster with Liu's?

```{r}
overlapping_GCFs <- long_bigscape %>%
  filter(author %in% c("Singleton", "Liu")) %>%
  group_by(GCF, cutoff) %>%
  summarise(from_singleton = sum(author == "Singleton"), 
            from_liu = sum(author == "Liu")) %>%
  filter(from_singleton > 0) %>%
  filter(from_liu > 0) 
  
overlapping_GCFs %>%
  group_by(cutoff) %>%
  summarise(number_of_overlapping_GCFs = n(), 
            bgcs_from_singleton = sum(from_singleton), 
            bgcs_from_liu = sum(from_liu), 
            total_bgcs = bgcs_from_singleton + bgcs_from_liu) 
  
```

Yes, there is some overlap. 60 BGCs overlap both studies, containing a total of 194 (4.58%) of our BGCs and 109 (5.19%) of theirs. 

# How much do the studies overlap with each other

```{r}
studies_plus_mibig <- c(levels(studies$author), "MIBiG")
bgc_count_mibig <- rbind(bgc_count, tibble(author = "MIBiG", total_bgcs = 1926))

find_overlap <- Vectorize(function(study1, study2, cutoff_value = 0.3){

 filter(long_bigscape, cutoff == cutoff_value) %>%
    mutate(author = if_else(
      str_detect(bgc_id, "^BGC") & is.na(author), "MIBiG", as.character(author)), 
      author = factor(author, levels = studies_plus_mibig))  %>%
    filter(author %in% c(study1, study2)) %>%
    group_by(GCF) %>%
    summarise(
      from_study1 = sum(author == study1), 
      from_study2 = sum(author == study2)) %>%
    filter(from_study1 > 0) %>%
    filter(from_study2 > 0)  %>%
    pull(from_study1) %>% sum()
})

overlap_matrix <- expand.grid(study1 = studies_plus_mibig,
                              study2 = studies_plus_mibig) %>%
  mutate(bgcs = find_overlap(study1, study2)) %>%
  mutate(bgcs = replace(bgcs, study1 == "MIBiG" & study2 == "MIBiG", 1926)) %>% 
  left_join(bgc_count_mibig, by = c("study1" = 'author')) %>%
  mutate(perc = as.double(bgcs/total_bgcs * 100), 
         across(starts_with('study'), 
                ~factor(.x, levels = studies_plus_mibig)))

overlap <-  ggplot(overlap_matrix,
                   aes(x = study1,
                       y = study2, 
                       fill = perc,
                       label = bgcs)) +
  geom_tile() + 
  geom_text() +
  scale_fill_gradient2(low = "steelblue4",
                       mid = "white", 
                       high = "firebrick", 
                       midpoint = 50) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.title = element_blank(),
        legend.position = "top", 
        legend.margin = margin(0,0,0,0))+
  guides(fill = guide_colorbar(title = "% of BGCs"))
  
overlap
```

# What about with other cutoff values?

```{r}
overlap_matrix_cutoffs <- 
  expand.grid(study1 = studies_plus_mibig, 
              study2 = studies_plus_mibig, 
              cutoff = c(0.3, 0.4, 0.5)) %>%
  mutate(bgcs = find_overlap(study1, study2, cutoff)) %>%
  mutate(bgcs = replace(bgcs, study1 == "MIBiG" & study2 == "MIBiG", 1926)) %>% 
  left_join(bgc_count_mibig, by = c("study1" = 'author')) %>%
  mutate(perc = as.double(bgcs/total_bgcs * 100), 
         across(starts_with('study'), 
                ~factor(.x, levels = studies_plus_mibig)), 
                cutoff = paste("cutoff =", cutoff))

overlap_faceted <- ggplot(overlap_matrix_cutoffs,
                   aes(x = study1,
                       y = study2, 
                       fill = perc,
                       label = bgcs)) +
  geom_tile() + 
  geom_text() +
  facet_grid(~cutoff) +
  scale_fill_gradient2(low = "steelblue4",
                       mid = "white", 
                       high = "firebrick", 
                       midpoint = 50) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.title = element_blank(),
        legend.position = "top", 
        legend.margin = margin(0,0,0,0), 
        panel.grid = element_blank(),
        axis.ticks = element_line(color = "black"))+
  guides(fill = guide_colorbar(title = "% of BGCs"))

ggsave("figures/overlap.svg", 
       overlap_faceted,
       width = 9, height = 4.3)


```

# How do these overlapping GCFs look like?

```{r}
nodes <- bgcs %>%
  filter(GCF_c0.3 %in% overlapping_GCFs$GCF) %>%
  filter(author %in% c("Singleton", "Liu")) %>%
  relocate(bgc_id, .before = 1) 

edges_df <- import_networks(bigscape_dir) %>%
  filter(if_all(starts_with("bgc_id"), ~.x %in% nodes$bgc_id)) 

graph <- graph_from_data_frame(edges_df, vertices = nodes, directed = FALSE)

vertex_attr(graph)$class <- factor(vertex_attr(graph)$class, levels = unique(my_classes))

layout <- layout_with_fr(graph, weights = edge_attr(graph)$squared_similarity)

ggraph(graph, layout = layout) +
  geom_edge_link(aes(alpha = squared_similarity)) +
  geom_node_point(aes(color = class, shape = author)) +
  scale_color_manual(values = my_colors, drop = FALSE) +
  #facet_nodes(~class) + 
  theme_classic() +
  theme(
   # legend.position = "none",
    axis.text = element_blank(), 
    axis.title = element_blank(), 
    axis.ticks = element_blank())

```

Modules

```{r}
modules <- data %>%
  map("modules") %>%
  bind_rows(.id = "study") %>%
 # filter(complete) %>%
  left_join(bind_rows(map(data, "regions"), .id = "study"),
            by = c("genome_id", "contig", "study"), 
            suffix = c("", "_region")) %>%
  # This could be a useful function
  mutate(.keep = "unused", across(
    starts_with("location"), list(
      start = ~str_match(.x, '\\[(.+?):')[,2],
      end = ~str_match(.x, ':(.+?)\\]')[,2]),
    .names = "{.col}_{.fn}")) %>%
  filter(location_start >= location_region_start) %>%
  filter(location_end <= location_region_end) %>%
  left_join(studies[,c("author", "reference")], by = c("study" = "reference")) %>%
  select(-contains("region"))

modules %>%
  mutate(product = map_chr(product, paste, collapse = " ")) %>%
  write_csv("tables/comparison_modules.csv")
```


```{r}
modular_products <- c("NRPS", "T1PKS", "transAT-PKS")
                
modules_df <- modules %>%
  filter(complete) %>%
  filter(map_lgl(product, ~as.logical(sum(modular_products %in% .x)))) %>% 
  mutate(product = map_chr(product, paste, collapse = ";")) %>%
  group_by(bgc_id, author) %>%
  summarise(n_modules = n()) %>%  
  #mutate(author = factor(study, levels = studies$author)) %>%
  arrange(author) %>%
  group_by(author) %>%
  mutate(n_bgcs = n(), 
         label = paste0(n_bgcs, " modular\nBGCs"), 
         label = factor(label, levels = unique(label))) 

modules_plot <- 
  ggplot(modules_df, aes(x = label, y = n_modules)) +
  geom_boxplot() +
  scale_y_log10(breaks = c(1,2,5,10,20)) +
  theme_minimal() +
  xlab("") +
  ylab(label = paste(
    "Number of complete PKS / NRPS modules in\nmodular BGCs (", 
    paste0(modular_products, collapse = " / "),
    ")")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5))

modules_plot
```
It seems that in the environment, multimodular BGCs are not as common as in isolates. 

```{r fig.height=7, fig.width=9}
modules_long_reads <- 
  modules_df %>%
  filter(author %in% studies[studies$long_reads == "long reads",]$author) %>%
  mutate(label = paste0(author, "\n(", label,  ")"), 
         label = factor(label, levels = unique(label))) %>%
  ggplot(aes(x = label, y = n_modules)) +
  geom_boxplot() +
  scale_y_log10(breaks = c(1,2,5,10,20)) +
  theme_minimal() +
  theme(axis.title = element_text(size = 8), 
        axis.line = element_line())+
  xlab("") +
  ylab(label = paste(
    "Number of complete PKS/NRPS\nmodules in modular BGCs\n(", 
    paste0(modular_products, collapse = " / "),
    ")")) 

modules_long_reads

```


```{r fig.height=7, fig.width=9}
panel_filepath <- "figures/comparison_panel.svg"

source("notebook/remove_text_length.R")

panel_2 <- 
  plot_grid(
    ncol = 2,
    plot_grid(
      ncol = 1, 
      contig_edge_2, 
      traffic_light,
      class_distribution_2, 
      rel_heights = c(7,4,7), 
      labels = c("A", "", "B")), 
    plot_grid(
      ncol = 1,
      overlap, 
      modules_long_reads, 
      rel_heights = 3:2, 
      labels = c("C", "D")),
    rel_widths = 3:2)

panel_2

panel_filepath <- "figures/comparison_panel.svg"

ggsave(panel_2, device = "svg", height = 7, width = 9, file = panel_filepath)
remove_text_length(panel_filepath)  
```
```{r}
long_bigscape %>%
  filter(cutoff == 0.3) %>%
  filter(author %in% c("Singleton", "Liu")) %>%
  group_by(GCF) %>%
  summarise(Singleton = sum(author == "Singleton"), 
            Liu = sum(author == "Liu")) %>%
  filter(Singleton > 0) %>%
  filter(Liu > 0) %>% nrow()

```

